---
title: "General Social Survey"
author: "Servio Maldonado"
date: "October 7th 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analylis are both categorical;hence we use bar graphs and tables to summarize the data, and a Chi-squared test for inference.

$H_O$: There is no relationship between gun ownership and weather one is in favor of capital punishment.

$H_A$: There is a relationship between people who own guns and people in favor of capital punishment.


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN+OWNGUN,data=mydata4, main='capital punishmnet and gun ownership',type='percent',xlab='capital punishment')
```

This barchart shows that people who own guns do in fact favor capital punishment more than the people who do not own guns.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1=xtabs(~CAPPUN+OWNGUN,data=mydata4)
rowPerc(tab1)
```

This table shows that there is no relationship between gun ownership and favoring or opposing capital punishment. The results are independent.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->



```{r}
chisqtestGC(tab1)
```
The p-vale for the chi sqare test is 0.02 < 0.05 so we reject the null hypotheis and conclude we have evidence to support the contention that gun ownership and wheather on favors capital punishment punishment are related.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->



Both variables in this analylis are both categorical;hence we use bar graphs and tables to summarize the data, and a Chi-squared test for inference.

$H_O$: There is no relationship between people who beleive in the Big Bang Theory and people who beleive in the afterlife.

$H_A$: There is a relationship between people who beleive in the Big Bang Theory and people who beleive in the afterlife.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~BIGBANG+POSTLIFE,data=mydata4,main='Bigbang and Postlife',type='percent',xlab='Big Bang')
```

This barchart shows that there is no corelation between the Big Bang theory beleivers and people who beleive in the afterlife.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
tab1 = xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(tab1)
```
This numerical table shows that there is no correlation between people who beleive in the Big Bang and their opinions of the afterlife, The results are independent of each other. 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisqtestGC(tab1)
```
The p-vale for the chi sqare test is 0.08 > 0.05 so we accept the null hypotheis and conclude that there is evidence to support the contention that people who beleive in the Big Bang Theory do beleive in the afterlife.


# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analylis are both categorical;hence we use bar graphs and tables to summarize the data, and a Chi-squared test for inference.

$H_O$: There is no relationship between between a persons race and their opinion on capital punishment.

$H_A$: There is a relationship between a persons race and their opinion on capital punishment.
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



```{r}

barchartGC(~RACE+CAPPUN, data=mydata4, main='Race and Capital Punishment', type='percent', xlab='Capital Punishment')
```
This barchart shows that there is a correlation between race and their stand on capital punishment.The results are related to each other. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
tab1 = xtabs(~RACE+CAPPUN, data=mydata4)
rowPerc(tab1)
```
This numerical table shows that there is a relation between the race of a person and whether they oppose or favor capital punishment. the results are related to each other.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```

The p-value for the chisq test is 9e-04 < 0.05  this suggest that there is a connection between a person's race and their stance on capital punishment.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Since the p- value for the chisq test is 9e-04 < 0.05 We fail to reject the null hypothesis and conclude that there is enough evidence to support that there is a relation between race and those who favor and oppose capital punishment.  

